Storage challenges in heterogeneous environments

### 2017-12-02 paper.tex
fix author list
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Noah Watkins, Michael Sevilla, Jeff LeFevre, Carlos Maltzahn \\
\affaddr{University of California, Santa Cruz} \\
\affaddr{\{nmwatkin,msevilla,jlefevre,carlosm\}@ucsc.edu}
}

@@

\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

Future extreme heterogeneous systems will be managed by runtimes that handle the
complexities of abstracting across systems with diverse sets of compute
elements, deep memory hierarchies, and asymmetrical networks.  The increased
role of runtimes in resource and execution management is a direct result of the
complexity involved in designing applications that are able to portably extract
performance from complex architectures.  Often the target of domain-specific
languages, such runtimes exploit application-specific semantics and use the
state of the current system (e.g., hardware, software, configuration, etc.) to
optimize all aspects of distributed execution. Runtimes such as Legion~\cite{bauer:sc12-legion}
expect to manage the entire compute infrastructure, but real-world constraints
require access to mass storage to be shared since it is used across applications
within organizations to relieve memory pressure, archive results, and provide
critical fault-tolerant capabilities. As a result, it is unlikely that in the
near-term the entire storage system hierarchy will be managed by a single
application runtime. Thus, the ability of application runtimes to optimize
execution is partly limited by the capabilities of I/O interfaces.

\begin{quote}
  \emph{Challenge}. Current storage architectures are not designed for extreme
  heterogeneity, especially in the context of persistent memories whose
  performance is limited code path length and shared-memory designs.
\end{quote}

Current I/O interfaces are treated like black boxes by runtime systems, limited
in use to basic read-write APIs (e.g., POSIX file I/O or middleware such as
HDF5). However, such interfaces are primarily designed for consumption by human
developers, rather than runtime systems.  In order to realize end-to-end
optimizations, I/O interfaces must be enhanced to expose storage capabilities
such as fine-grained control over data durability, horizontal and vertical data
locality, and semantics that affect locking and serialization. Runtime systems
may take advantage of these capabitilies in diverse, application-specific ways.
But as a shared resource, storage systems cannot expose exclusive control to an
application runtime. In fact, important optimizations in storage systems depend
on managing access by understanding behavior across application boundaries.

\begin{quote}
  \emph{Challenge}. The design of adaptive storage I/O interfaces that safely
  provide two-way communication of with application runtimes the nature of
  workflows, hardware, data, and access semantics. Highly dynamic interfaces to shared resources
  present many optimization opporunities, but introduce difficult challenges for
  managing access.
\end{quote}

@@

\subsection{Interface design}

Current I/O interfaces are in large part treated as black boxes that offer
limited control in the form of \emph{magic numbers} or coarse-grain hints. With
few exceptions such as I/O sequentiality and alignment parameters, existing
interfaces do not allow applications to express intent or expose low-level system state
or configurability. In order to realize {\it end-to-end} application I/O optmizations,
both application runtime and storage system will need to be co-designed such that runtime
adapation of behavior based on data structure is possible.

For instance, HPC applications that use pattern-based data models and want to
optimize for checkpoint latency should be able to reliqunish control over low-level
data placement to the storage system, and in turn be able to adapatively provide
subsets of checkpoint data based on the current state of the storage system.

Historically the primary I/O interfaces have been object, block, and POSIX file
I/O. However, the availability of high-performance open-source storage systems
are allowing exploration of new storage interfaces without the fear of vendor
lock-in.  Recent work has explored the adaptability of existing storage systems
to support a range of new I/O interfaces, but an open question is how such
interfaces can efficently and reliability be constructed. Integrate KV-Drive and
LDFI into discussion.

@@

Both HPC and cloud-based systems are seeing rapid advancements towards
heterogeneity. For instance, Amazon Web Services offers a large array of node
and network capabilities including FPGA and GPU co-processors. Moreover, AWS
offers an HPC service for users to compose their own HPC systems Emerging
storage technologies such as NVMe and persistent memories are offered by cloud
providers and will be integrated in future HPC systems, but even existing
storage systems are struggling to exploit their full
performance~\cite{xu:fast16-nova}. In fact, the increased performance of
next-generation media and networks suggests that future systems may require a
CPU per storage device~\cite{samuels:oss16}. And in order to support a dynamic
range of interfaces, internal sub-systems will need to be generic and re-usable
to support the diverse of demands of runtime systems.

Achieving end-to-end performance optimization will require new techniques for
verticaly integrating across application, runtime, and storage systems. The
architectural challenges of achieving low-latency I/O for persistent memories
requirements are significant. These media benefit from shared-nothing designs
that minimize context switches and \emph{cycles}, thus one-size-fits-all code
paths will be replaced by many specialized paths, potentially constructed
dynamically, that each must still support correct consistency semantics across
interfaces.

The incredible advancement of container and virtualization technologies,
dependence on domain-specific languages, and reliance on open-source software
for critical support libraries, are increasing the demand for polyglot
deployments. However, technologies for cross-language communication offer
notoriously poor performance (e.g. JNI) and will introduce unacceptable
overheads for systems with next-generation memories. New runtime systems that
rely on advanced JIT technologies are successfully being used to dynamically
perform end-to-end code path optimization across language boundaries (Graal,
Truffle, SQLiteJIT).

@@

The data-intensive nature of both cloud and HPC applications will put extreme
pressure on shared storage systems. Even in the most constrained cases today,
efficently sharing systems is a significant challenge. Handling access to a
shared storage systems supporting a wide range of dynamic interfaces, code
paths, and new architectures that favor less shared state will require
rethinking how resource management is handled.

@@

As with the other research directions, operational policies reach beyond just
HPC. Providing performance guarantees for all resources, not just storage, is
important for *aaS architectures that are shared among users, so lessons can be
gleaned from cloud research about isolation and scheduling, such as Mesos and
Kubernetes. Furthermore, the cloud community has done a good job addressing
heterogeneity when using these techniques behind user-facing APIs without
jeopardizing the semantics.

>>>>

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Noah Watkins, Michael A. Sevilla, Jeff LeFevre, Carlos Maltzahn \\
\affaddr{University of California, Santa Cruz} \\
\affaddr{\{nmwatkin,msevilla,jlefevre,carlosm\}@ucsc.edu}
}

@@

\date{\today}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

Future extreme heterogeneous systems will need to be managed by runtimes that handle the
complexities of abstracting across systems with diverse sets of compute
elements, deep memory hierarchies, and asymmetrical networks.  The increased
role of runtimes in resource and execution management is a direct result of the
complexity involved in designing portable applications that are able to extract
performance from complex architectures.  Often the target of domain-specific
languages, such runtimes exploit application-specific semantics and use the
state of the current system (e.g., hardware, software, configuration, etc.) to
optimize all aspects of distributed execution. Runtimes such as Legion~\cite{bauer:sc12-legion}
expect to manage the entire compute infrastructure, but real-world constraints
require access to mass storage to be shared since it is used across applications
and organizations to relieve memory pressure, archive results, and provide
critical fault-tolerant capabilities. As a result, it is unlikely that in the
near-term the entire storage system hierarchy will be managed by a single
application runtime. Thus, the ability of application runtimes to optimize
execution is partly limited by the capabilities of I/O interfaces and the
opportunities by which the storage system can provide cross-application optimizations.

\begin{quote}
	\emph{Challenge}. Current storage architectures are not designed for extreme
heterogeneity, in terms of both hardware as well as users, especially
in the context of persistent memories whose performance is limited by code path
length and shared-memory designs.
\end{quote}

Current I/O interfaces are treated like black boxes by runtime systems, limited
in use to basic read-write APIs (e.g., POSIX file I/O or middleware such as
HDF5). However, such interfaces are primarily designed for consumption by human
developers, rather than runtime systems.  In order to realize end-to-end
optimizations, I/O interfaces must be enhanced to expose storage capabilities
such as fine-grained control over data durability, horizontal and vertical data
locality, and semantics that affect locking and serialization. Runtime systems
must be able to take advantage of these capabilities in diverse, application-specific ways.
But as a shared resource, storage systems cannot expose exclusive control to an
application runtime. In fact, important optimizations in storage systems depend
on managing access by understanding behavior across application and process boundaries.

\begin{quote}
  \emph{Challenge}. The design of adaptive storage I/O interfaces that safely
  provide two-way communication with application runtimes regarding the nature of
  workflows, hardware, data, and access semantics. Highly dynamic interfaces to shared resources
  present many optimization opportunities, but introduce difficult challenges for
  managing access.
\end{quote}

These challenges are a direct result of the disruptive nature of
next-generation storage media that are recognized by the storage industry as
drivers of new storage system architectures~\cite{samuels:oss16}.

@@

\newpage
\subsection{Interface design}

Current I/O interfaces are in large part treated as black boxes that offer
limited control in the form of \emph{magic numbers} or coarse-grain hints. With
few exceptions such as I/O sequentiality and alignment parameters, existing
interfaces do not allow applications to express intent or expose low-level system state
or configurability. In order to realize {\it end-to-end} application I/O optimizations,
both application runtime and storage system will need to be co-designed such that runtime
adaptation of behavior is possible.

For instance, HPC applications that use pattern-based data models should be able
to relinquish control over low-level data placement to the storage system, and
in turn be able to adaptively provide subsets of checkpoint data based on the
current state of the storage system. This push-pull model allows both
application and storage system to cooperatively adjust behavior.

Historically the primary I/O interfaces have been object, block, and POSIX file
I/O. However, the availability of high-performance open-source storage systems
are allowing exploration of new storage interfaces without the fear of vendor
lock-in.  Recent work has explored the adaptability of existing storage systems
to support a range of new I/O interfaces, but an open question is how such
interfaces can efficiently and reliability be constructed.
%TODO Integrate KV-Drive and
%LDFI into discussion.

@@

Both HPC and cloud-based systems are seeing rapid advancements towards
heterogeneity. For instance, Amazon Web Services offers a large array of node
and network capabilities including FPGA and GPU co-processors. Moreover, AWS
offers an HPC service for users to compose their own HPC systems. Emerging
storage technologies such as NVMe and persistent memories are offered by cloud
providers and will be integrated in future HPC systems, but even existing
storage systems are struggling to exploit their full
performance~\cite{xu:fast16-nova}. In fact, the increased performance of
next-generation media and networks suggests that future systems may require a
CPU per storage device~\cite{samuels:oss16}. And in order to support a dynamic
range of interfaces, internal sub-systems will need to be generic and re-usable
to support the diverse demands of runtime systems.

Achieving end-to-end performance optimization will require new techniques for
vertically integrating across application, runtime, and storage systems. The
architectural challenges of achieving low-latency I/O for persistent memories
requirements are significant. These media benefit from shared-nothing designs
that minimize context switches and \emph{cycles}, thus one-size-fits-all code
paths will be replaced by many specialized paths, potentially constructed
dynamically, that each must still support correct consistency semantics across
interfaces.

The incredible advancement of container and virtualization technologies,
dependence on domain-specific languages, and reliance on open-source software
for critical support libraries, are increasing the demand for polyglot
deployments. However, technologies for cross-language communication offer
notoriously poor performance (e.g. JNI) and will introduce unacceptable
overheads for systems with next-generation memories. New runtime systems that
rely on advanced JIT technologies are successfully being used to dynamically
perform end-to-end code path optimization across language
boundaries~\cite{wurt,friedrich:ecoop16}.

@@

The data-intensive nature of both cloud and HPC applications will put extreme
pressure on shared storage systems. Even in the most constrained cases today,
efficiently sharing systems is a significant challenge. Handling access to a
shared storage systems supporting a wide range of dynamic interfaces, code
paths, and new architectures that favor less shared state will require
rethinking how resource management is handled.

@@

As with the other research directions, operational policies reach beyond just
HPC. Providing performance guarantees for all resources, not just storage, is
important for SaaS architectures that are shared among users, so lessons can be
gleaned from cloud research about isolation and scheduling, such as Mesos and
Kubernetes. Furthermore, the cloud community has done a good job addressing
heterogeneity when using these techniques behind user-facing APIs without
jeopardizing the semantics.
